---
title: "Retrieve QW Data"
author: "David J Holtschlag"
date: " `r format( Sys.Date(), '%A %B %d, %Y')` "
output: html_document
---

```{r setup, include=FALSE}
library(dataRetrieval)
library(tidyverse)
fig_no <- 0
knitr::opts_chunk$set(echo = FALSE)
```

## Retrieval of Water Quality Summaries for 04101500 St. Joseph River at Niles, MI and 04165500 Clinton River at Mt. Clemens, MI 

```{r get_data_summary}

# 04165500 Clinton River at Mt. Clemens
if(file.exists("site/04165500/qw_summary_04165500.RData")){
  load('site/04165500/qw_summary_04165500.RData')
}else {
  qw_summary_04165500      <- whatNWISdata( siteNumbers = '04165500', service = 'qw' )
  save(qw_summary_04165500, file = "site/04165500/qw_summary_04165500.RData")
}

# 04101500 St. Joseph River at Niles, MI
if(file.exists("site/04101500/qw_summary_04101500.RData")){
  load("site/04101500/qw_summary_04101500.RData")
}else {
  qw_summary_04101500      <- whatNWISdata( siteNumbers = '04101500', service = 'qw' )
  save(qw_summary_04101500, file = "site/04101500/qw_summary_04101500.RData")
}

# Read parameter codes downloaded from web url 
# https://nwis.waterdata.usgs.gov/nwis/pmcodes/pmcodes?radio_pm_search=param_group&pm_group=All+--+include+all+parameter+groups&pm_search=&casrn_search=&srsname_search=&format=rdb&show=parameter_group_nm&show=parameter_nm&show=casrn&show=srsname&show=parameter_units

pmcodes <- read.csv(file = 'data/pmcodes.txt', sep = '\t', header = TRUE,
                    comment.char = "#", colClasses = c('character'))

# Get indices of sorted parameter codes 
tmp_ndx <- order(pmcodes$parameter_cd)
# Re-order pmcodes to sorted values
pmcodes <- pmcodes[tmp_ndx, ]


```

## Inital Processing of Data Summaries to find High Frequency (>200) Samples

```{r init_process}

# Find parameters that have been sampled 200 or more times and are not missing para_cd
qw_summary_04165500_200  <- qw_summary_04165500[ which( qw_summary_04165500$count_nu >= 200 & !is.na(qw_summary_04165500$parm_cd)),]

# Find indices of parameter codes that have been sampled in all parameter codes
ndx_pmcodes <- which(pmcodes$parameter_cd %in% qw_summary_04165500_200$parm_cd)

# Populate fields with descriptive parameter info
qw_summary_04165500_200$group_nm <- pmcodes$parameter_group_nm[ndx_pmcodes]
qw_summary_04165500_200$parm_nm  <- pmcodes$parameter_nm[ndx_pmcodes]
qw_summary_04165500_200$srsname  <- pmcodes$srsname[ndx_pmcodes]

# Keep !'informational' parameter_group_nm
ndx_keep <- which(qw_summary_04165500_200$group_nm != 'Information')
qw_summary_04165500_200 <- qw_summary_04165500_200[ ndx_keep, ]


```

## Retrieve Selected Water Quality Parameters for Field Samples 


Angela Brennan suggests that either turbidity or Specific conductance might be used to estimate SSC or Ortho-P

```{r get_qw_data} 

site_ids    <- c('04101500', '04165500')
# Parameters:   SSC      Ortho-P  SC        Flow      Stage   H20_temp
parameterCd <- c('80154', '00671', '00095', '00061', '00065', '00010')

supp.labs   <- c('Suspended Sediment Concentration', 'Orthophosphate', 'Specific Conductance',
                'Flow', 'Stage', 'H20 Temperature')

names(supp.labs) <- c('80154', '00671', '00095', '00061', '00065', '00010')

if( file.exists("data/NWISqwData.RData")){
  load("data/NWISqwData.RData")
} else{
rawNWISqwData          <- readNWISqw(site_ids,parameterCd,reshape=FALSE)
NWISqwData <- rawNWISqwData %>% 
  filter( samp_type_cd == '9' & medium_cd == 'WS' & sample_dt > as.Date('2010-10-01') )
  save(NWISqwData, file = "data/NWISqwData.RData")
  rm(rawNWISqwData)
}
```

### Retrieve Unit Value Water-Quality Data and Compute Daily Values'

This code uses the dataRetrieval package to access unit value water-quality data
from 04101500 St. Joseph River at Niles, MI, from April 1, 2011 to September 30,
2020. The unit data were generally obtained at 15-minute intervals forming a
data set with 329,058 values on five water quality parameters along with flow
and gage height (stage) data. Most unit GH value were missing. Daily means were
computed from the unit values resulting in 3434 daily values. Means were
computed as long as there was at least on measured unit value.

```{r uv_dataRetrieval, eval = FALSE, fig.height = 20, fig.width = 9}

# Retrieve Unit Value Data
if( file.exists("data/rawNWISqw_UnitData.RData")){
  load("data/rawNWISqwData.RData")
} else{
  # 'uv_temp_water, 'uv_flow', 'uv_stage', 'uv_SC', 'uv_DO', 'uv_pH', 'uv_turbidity'
  uv_parm_cd <- c('00010', '00060', '00065', '00095', '00300', '00400', '63680')
  rawNWISqw_UnitData  <- readNWISuv(siteNumbers = '04101500', parameterCd = uv_parm_cd,
                                  startDate = '2011-04-01', endDate = '2020-09-30')
  rawNWISqw_UnitData <- renameNWISColumns(rawNWISqw_UnitData)
  NWISnames   <- colnames(rawNWISqw_UnitData)
  newNames    <- str_replace(NWISnames, 'Inst', 'uv')
  colnames(rawNWISqw_UnitData) <- newNames
  save(rawNWISqw_UnitData, file = 'data/rawNWISqw_UnitData.RData')
}

# Aggregate unit to daily values, which reduces n from 329058 to 3434
NWISqw_DayWide <- rawNWISqw_UnitData %>% 
  filter( as.Date(dateTime) > as.Date('2011-04-01')) %>% 
  group_by( as.Date(dateTime) ) %>% 
  summarise(Wtemp_day    = mean(Wtemp_uv,    na.rm = TRUE),
            Flow_day     = mean(Flow_uv,     na.rm = TRUE),
            GH_day       = mean(GH_uv,       na.rm = TRUE),
            SpecCond_day = mean(SpecCond_uv, na.rm = TRUE),
            DO_day       = mean(DO_uv,       na.rm = TRUE),
            pH_day       = mean(pH_uv,       na.rm = TRUE),
            Turb_day     = mean(Turb_uv,     na.rm = TRUE))
            # SpecCond_min = min(SpecCond_uv,  na.rm = TRUE),
            # SpecCond_max = max(SpecCond_uv,  na.rm = TRUE))
colnames(NWISqw_DayWide)[1] <- 'monitor_dt'

save(NWISqw_DayWide, file = 'data/NWISqw_DayWide.RData')

NWISqw_DayLong <- NWISqw_DayWide %>% 
  pivot_longer(cols = Wtemp_day:Turb_day) 

colnames(NWISqw_DayLong) <- c('Date', 'parameter', 'daily_mean')
save(NWISqw_DayLong, file = 'data/NWISqw_DayLong.RData')

```


### Plot Daily Means of Continuous Data 

```{r plot_daily_means, fig.height=20, fig.width = 9}

load("data/NWISqw_DayLong.RData")

fig_no <- fig_no + 1
NWISqw_DayLong %>% 
  ggplot( aes(x = Date, y = daily_mean, group = 'parameter')) +
  geom_point( size = 0.5) +
  facet_wrap( ~ parameter, ncol = 1, scales = 'free_y') +
  theme_bw() +
  labs(title = paste0('Figure ', fig_no, '. Time Series of Continuous Parameters at 04101500 St. Joseph River at Niles, MI'))
  


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### QW time series at 04101500 St. Joseph River at Niles, Michigan

```{r plot_series_1015, fig.height= 20, fig.width = 9 }

new.lab <- as_labeller(c('80154' = '80154 Suspended Sediment Concentration', '00671' = '00671 Orthophosphate', 
                         '00095' = '00095 Specific Conductance', '00061' = '00061 Flow', '00065' = '00065 Stage',
                         '00010' = '00010 Water Temperature'))

NWISqw1015 <- NWISqwData %>% filter( site_no == '04101500' ) 

fig_no <- fig_no + 1

NWISqw1015 %>% 
  select(sample_dt, parm_cd, result_va) %>% 
  ggplot( aes( x = sample_dt, y = result_va, group = parm_cd)) +
  geom_point() +
  facet_wrap(~parm_cd, ncol = 1, scales = 'free_y', labeller = new.lab ) +
  theme_bw() +
  labs(title = paste0('Figure ',fig_no,'. Selected Water Quality Parameters at 04101500 St. Joseph River at Niles, MI'))

```

### Relation between specific conductance and suspended sediment concentration at 04101500

```{r plot_SSC_relation_1015, fig.height = 6, fig.width = 8}
 NWISqw <- NWISqw1015[, c('site_no', 'sample_dt', 'parm_cd', 'result_va')] %>% 
  dplyr::filter( (parm_cd == '00095' | parm_cd == '80154'))  %>% 
  mutate(date_parm = paste(sample_dt, parm_cd, sep = '_')) %>% 
  group_by(date_parm) %>% summarize(result_mean = mean(result_va, na.rm = T), result_n = n(),
                                    result_change  = (max(result_va) - min(result_va) )/result_mean * 100.) %>% 
  select(date_parm, result_mean, result_n, result_change) %>% 
  mutate(sample_dt = as.Date( substr(date_parm, 1, 10) ),
         parm_cd   = as.character( substr(date_parm, 12, 16))) %>% 
  select(sample_dt, parm_cd, result_n, result_mean, result_change) %>% 
  pivot_wider( names_from = parm_cd, values_from = c( result_mean, result_change, result_n))
  
fig_no <- fig_no + 1

NWISqw %>% 
  mutate(change = pmax(result_change_00095,   result_change_80154) ) %>% 
  ggplot( aes( x =     result_mean_00095, y = result_mean_80154, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_80154 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'sqrt') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Specific Conduction 00095') +
  ylab('Suspended Sediment Concentration 80154') +
  labs( title = paste0('Figure ', fig_no, ' Relation between specific conduction and suspended sediment concentration \n at 04101500 St. Joseph River at Niles, Michigan.'))

```


### Relation between specific conductance and orthophosphate concentration at 04101500

```{r plot_PO4_relation_1015, fig.height = 6, fig.width = 8}

NWISqw <- NWISqw1015[, c('site_no', 'sample_dt', 'parm_cd', 'result_va')] %>% 
  dplyr::filter( (parm_cd == '00095' | parm_cd == '00671' | parm_cd == '00061') ) %>% 
  mutate(date_parm = paste(sample_dt, parm_cd, sep = '_')) %>% 
  group_by(date_parm) %>% summarize(result_mean = mean(result_va, na.rm = T), result_n = n(),
                                    result_change  = (max(result_va) - min(result_va) )/result_mean * 100.) %>% 
  select(date_parm, result_mean, result_n, result_change) %>% 
  mutate(sample_dt = as.Date( substr(date_parm, 1, 10) ),
         parm_cd   = as.character( substr(date_parm, 12, 16))) %>% 
  select(sample_dt, parm_cd, result_n, result_mean, result_change) %>% 
  pivot_wider( names_from = parm_cd, values_from = c( result_mean, result_change, result_n))

fig_no <- fig_no + 1
NWISqw %>% 
  mutate(change = pmax(result_change_00095,   result_change_00671) ) %>% 
  ggplot( aes( x =     result_mean_00095, y = result_mean_00671, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_00671 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'sqrt') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Specific Conduction 00095') +
  ylab('Orthophosphate 00671') +
  labs( title = paste0('Figure ', fig_no, '. Relation between specific conduction and orthophosphate concentration \n at 04101500 St. Joseph River at Niles, Michigan.'))

fig_no <- fig_no + 1
NWISqw %>% 
  mutate(change = pmax(result_change_00061,   result_change_00671) ) %>% 
  ggplot( aes( x =     result_mean_00061, y = result_mean_00671, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_00671 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'log10') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Streamflow 00061') +
  ylab('Orthophosphate 00671') +
  labs( title = paste0('Figure ', fig_no, '. Relation between streamflow and orthophosphate concentration \n at 04101500 St. Joseph River at Niles, Michigan.'))


```

## QW time series at 04165500 Clinton River at Mt. Clemens, Michigan

```{r plot_series_1655, fig.height= 20, fig.width = 9}


NWISqw1655 <- NWISqwData %>% filter( site_no == '04165500' ) 

fig_no <- fig_no + 1
NWISqw1655 %>% 
  select(sample_dt, parm_cd, result_va) %>% 
  ggplot( aes( x = sample_dt, y = result_va, group = parm_cd)) +
  geom_point() +
  facet_wrap(~parm_cd, ncol = 1, scales = 'free_y', labeller = new.lab  ) +
  theme_bw() +
  labs(title = paste0('Figure ', fig_no,'. Selected Water Quality Parameters at 04165500 Clinton River at Mt. Clemens, MI'))

```

### Relation between specific conductance and suspended sediment concentration at 04165500

```{r plot_SSC_relation_1665, fig.height = 6, fig.width = 8}

NWISqw <- NWISqw1655[, c('site_no', 'sample_dt', 'parm_cd', 'result_va')] %>% 
  dplyr::filter( (parm_cd == '00095' | parm_cd == '80154' | parm_cd == '00061') ) %>% 
  mutate(date_parm = paste(sample_dt, parm_cd, sep = '_')) %>% 
  group_by(date_parm) %>% summarize(result_mean = mean(result_va, na.rm = T), result_n = n(),
                                    result_change  = (max(result_va) - min(result_va) )/result_mean * 100.) %>% 
  select(date_parm, result_mean, result_n, result_change) %>% 
  mutate(sample_dt = as.Date( substr(date_parm, 1, 10) ),
         parm_cd   = as.character( substr(date_parm, 12, 16))) %>% 
  select(sample_dt, parm_cd, result_n, result_mean, result_change) %>% 
  pivot_wider( names_from = parm_cd, values_from = c( result_mean, result_change, result_n))
  
fig_no <- fig_no + 1
NWISqw %>% 
  mutate(change = pmax(result_change_00095,   result_change_80154) ) %>% 
  ggplot( aes( x =     result_mean_00095, y = result_mean_80154, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_80154 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'sqrt') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Specific Conduction 00095') +
  ylab('Suspended Sediment Concentration 80154') +
  labs( title = paste0('Figure ', fig_no, '. Relation between specific conduction and suspended sediment concentration \n at 04165500 Clinton River at Mt. Clemens, Michigan.'))

fig_no <- fig_no + 1
NWISqw %>% 
  mutate(change = pmax(result_change_00061,   result_change_80154) ) %>% 
  ggplot( aes( x =     result_mean_00061, y = result_mean_80154, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_80154 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'sqrt') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Streamflow 00061') +
  ylab('Suspended Sediment Concentration 80154') +
  labs( title = paste0('Figure ', fig_no, '. Relation between streamflow and suspended sediment concentration \n at 04165500 Clinton River at Mt. Clemens, Michigan.'))


```

### Relation between specific conductance and suspended sediment concentration at 04165500

```{r plot_PO4_relation_1665, fig.height = 6, fig.width = 8}

NWISqw <- NWISqw1655[, c('site_no', 'sample_dt', 'parm_cd', 'result_va')] %>% 
  dplyr::filter( (parm_cd == '00095' | parm_cd == '00671' | parm_cd == '00061') ) %>% 
  mutate(date_parm = paste(sample_dt, parm_cd, sep = '_')) %>% 
  group_by(date_parm) %>% summarize(result_mean = mean(result_va, na.rm = T), result_n = n(),
                                    result_change  = (max(result_va) - min(result_va) )/result_mean * 100.) %>% 
  select(date_parm, result_mean, result_n, result_change) %>% 
  mutate(sample_dt = as.Date( substr(date_parm, 1, 10) ),
         parm_cd   = as.character( substr(date_parm, 12, 16))) %>% 
  select(sample_dt, parm_cd, result_n, result_mean, result_change) %>% 
  pivot_wider( names_from = parm_cd, values_from = c( result_mean, result_change, result_n))
  
fig_no <- fig_no + 1

NWISqw %>% 
  mutate(change = pmax(result_change_00095,   result_change_00671) ) %>% 
  ggplot( aes( x =     result_mean_00095, y = result_mean_00671, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_00671 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'sqrt') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Specific Conduction 00095') +
  ylab('Orthophosphate 00671') +
  labs( title = paste0('Figure ', fig_no, '. Relation between specific conduction and orthophosphate concentration \n at 04165500 Clinton River at Mt. Clemens, Michigan.'))

fig_no <- fig_no + 1

NWISqw %>% 
  mutate(change = pmax(result_change_00061,   result_change_00671) ) %>% 
  ggplot( aes( x =     result_mean_00061, y = result_mean_00671, color = change )) +
  geom_point(size = 1 / NWISqw$result_n_00671 ) +
  scale_y_continuous( trans = 'sqrt') +
  scale_x_continuous( trans = 'sqrt') +
  theme_bw() +
  theme( legend.position = 'bottom') +
  xlab('Streamflow 00061') +
  ylab('Orthophosphate 00671') +
  labs( title = paste0('Figure ', fig_no, '. Relation between streamflow and orthophosphate concentration \n at 04165500 Clinton River at Mt. Clemens, Michigan.'))


```

### Compare probability densities at 04101500 and 01165500

```{r compare_densities, fig.width = 8, fig.height = 5}
NWISqw <- NWISqwData %>% filter( parm_cd == '80154') %>% 
  select( site_no, sample_dt, parm_cd, result_va)

fig_no <- fig_no + 1

NWISqw %>% 
  ggplot( aes( log10(result_va )) ) +
  geom_density( aes(group = site_no, color = site_no ) ) +
  theme_bw() +
  xlab('Log10 Suspended Sediment Concentration') +
  ylab('Probability Density') +
  labs( title = paste0('Figure ', fig_no, '. Probability densities of Suspended Sediment Concentrations at \n04101500 St. Joseph River at Niles and 04165500 Clinton River at Mt. Clemens'))
  
NWISqw <- NWISqwData %>% filter( parm_cd == '00671') %>% 
  select( site_no, sample_dt, parm_cd, result_va)

fig_no <- fig_no + 1
NWISqw %>% 
  ggplot( aes( log10(result_va )) ) +
  geom_density( aes(group = site_no, color = site_no ) ) +
  theme_bw() +
  xlab('Log10 Orthophosphate Concentration') +
  ylab('Probability Density') +
  labs( title = paste0('Figure ', fig_no, '. Probability densities of Orthophosphate Concentrations at \n04101500 St. Joseph River at Niles and 04165500 Clinton River at Mt. Clemens'))

```


### Compare discrete samples with continuous monitoring data

Using water-quality data from site 04101500 St. Joseph River at Niles, compare discrete and continuous daily mean of specific conductance 

```{r merge_discrete_daily_means}

# Extracts a subset of variables and computes a mean for sample_dt & parm_cd dups
NWISqwLong <- NWISqw1015[, c('site_no', 'sample_dt', 'parm_cd', 'result_va')] %>% 
  group_by( sample_dt, parm_cd) %>% 
  summarize(result_mean = mean(result_va, na.rm = T), result_n = n()) 


NWISqwWide <- NWISqwLong %>% 
  dplyr::select( sample_dt, parm_cd, result_mean ) %>% 
  pivot_wider( names_from = parm_cd, values_from = result_mean )

df <- data.frame( SpecCond_sample = c(332, 332), SpecCond_day = c(649, 664) )

load('data/NWISqw_DayWide.RData')

SC_sample_cont <- left_join(NWISqwWide[, c('sample_dt', '00095')], NWISqw_DayWide[,c('monitor_dt','SpecCond_day')],
                            by = c('sample_dt' = 'monitor_dt'))
colnames(SC_sample_cont)[2] <- 'SpecCond_sample'

fig_no <- fig_no + 1
SC_sample_cont %>% 
  ggplot( aes( x = SpecCond_sample, y = SpecCond_day) ) +
  geom_point() +
  geom_segment(aes( x = df$SpecCond_sample[1], xend = df$SpecCond_sample[2],
                    y = df$SpecCond_day[1]   , yend = df$SpecCond_day[2]), color = 'red') +
  geom_abline( intercept = 0, slope = 1) +
  theme_bw() +
  xlab('Discrete measurement of Specific Conductance ') +
  ylab('Daily mean of 15-minute specific conductance values') +
  labs(title = paste0('Figure ', fig_no, '. Relation between discrete and daily mean of continuous specific \nconductance nvalues at 04101500 St. Joseph River at Niles, MI'))




```